---
title: "RETP H1a, H1b, 2, 3"
author: "Frederieke"
date: "2026-04-05"
output: html_document
---

-----------------------Analysis Hypotheses 1a,1b,2,3 ---------------------------

This markdown features the analysis pipeline for model 1, used to test hypothesis 1a, 2, and 3.

--------------------------------------------------------------------------------



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading packages

```{r}
pacman::p_load(tidyverse, rlang, tidymodels, gridExtra, brms, cmdstanr, truncnorm, posterior, bayesplot)
```


### Read in dataframe
```{r}
dt <- read_csv("C:/.../Analysis/all_data.csv")
```


##################### Modeling##############################################


```{r}
#defining model

m4 <- bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type + 
           (0 + condition:orientation:aoi_type:media_type | ID) + 
           (0 + condition:orientation:aoi_type:media_type | object_id))


```

############################## Priors

### Beginning by getting some estimates from the data

```{r}
dt %>%
  group_by(aoi_type, orientation) %>%
  summarise(
    mean_dwell = mean(trans_dwell),
    sd_dwell   = sd(trans_dwell),
    .groups = "drop"
  )
```


############### Model 4 execution ##############################################

```{r}
get_prior(
  bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type + 
           (0 + condition:orientation:aoi_type:media_type | ID) + 
           (0 + condition:orientation:aoi_type:media_type | object_id),
     phi ~ 1),
  data   = dt,
  family = Beta()
)
```

```{r}
#mainly using weakly informative priors
pr4 <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor")
)
```

#### Run priors

```{r}
m4_prior <-
  brm(
    m4,
    data = dt,
    prior = pr4,
    family = Beta(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "real_m4_prior",
    warmup = 500,
    iter = 1000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```

### Prior checks

```{r}
pp4 <- pp_check(m4_prior, ndraws = 100)
```

```{r}
#visualise prior checks
pp4 + xlim(-1, 2)
```


### Fitting the posterior model

```{r}
m4_pos <- brm(
    m4,
    data = dt,
    family = Beta(link = "logit"),
    prior = pr4,
    sample_prior = T,
    iter = 4000,
    warmup = 2000,
    cores = 2,
    chains = 4,
    seed = 1012,
    backend = "cmdstanr",
    threads = threading(4),
    file = "real_m4_pos",
    control = list(
      adapt_delta = 0.95,
      max_treedepth = 15 ))


```

```{r}
#long and convoluted summary
summary(m4_pos)
```




############# Prior Posterior Update Checks/Convergence############################


```{r}
#check bulk and tail, rhat diagnsotics separately

draws <- as_draws(m4_pos)

# Get all diagnostics in one go
diagnostics <- summarise_draws(draws, rhat, ess_bulk, ess_tail)
```



### Example Caterpillars
```{r}
#given the complexity, only visualise a few select ones
samples <- readRDS("real_m4_pos.rds")
```


```{r}
#fixed labelled, orientation 0, art aoi, art  object
trace1 <- mcmc_trace(as.array(samples), pars = "b_conditionlabelled:orientation0:aoi_typea:media_typea") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()

#fixed labelled, orientation 0, art aoi, tool object
trace2 <- mcmc_trace(as.array(samples), pars = "b_conditionlabelled:orientation0:aoi_typea:media_typet") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()

#fixed , labelled, orientation 0, tool aoi, tool object
trace3 <- mcmc_trace(as.array(samples), pars = "b_conditionlabelled:orientation0:aoi_typet:media_typet") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()

#fixed unlabelled, orientation 0, art aoi, art object
trace4 <- mcmc_trace(as.array(samples), pars = "b_conditionunlabelled:orientation0:aoi_typea:media_typea") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()


#Random effect
trace5 <- mcmc_trace(as.array(samples), pars = "sd_ID__conditionlabelled:orientation0:aoi_typea:media_typea") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()

#random object, labelled
trace6 <- mcmc_trace(as.array(samples), pars = "sd_object_id__conditionlabelled:orientation0:aoi_typea:media_typea") +
  ggtitle("Trace Plot for Intercept") +
  theme_bw()

```


#visualise traceplots
```{r}
trace1

trace2

trace3

trace4

trace5

trace6

```



```{r}
draws_df <- as_draws_df(m4_pos)
```


```{r}
#grab names
names(draws_df)[grepl("prior", names(draws_df))]
```





### Prior posterior plots
```{r}
#Again only examples due to model complexity

#Remember, Prior columns are automatically named with prior_
pp1 <- ggplot(draws_df) +
  geom_density(aes(x = `b_conditionlabelled:orientation0:aoi_typea:media_typea`), fill = "blue", alpha = 0.3) +
  geom_density(aes(x = `prior_b`), fill = "red", alpha = 0.3) +
  labs(title = "Prior vs Posterior",
       x     = "Effect (log-odds scale)",
       y     = "Density") +
  xlim(c(-2, 6)) +
  theme_bw()

pp2 <- ggplot(draws_df) +
  geom_density(aes(`b_conditionlabelled:orientation0:aoi_typet:media_typet`), fill = "blue", alpha = 0.3) +
  geom_density(aes(`prior_b`), fill = "red", alpha = 0.3) +
  labs(title = "Prior vs Posterior",
       x     = "Effect (log-odds scale)",
       y     = "Density") +
  xlim(c(-2, 6)) +
  theme_bw()

pp3 <- ggplot(draws_df) +
  geom_density(aes(`b_conditionunlabelled:orientation0:aoi_typea:media_typea`), fill = "blue", alpha = 0.3) +
  geom_density(aes(`prior_b`), fill = "red", alpha = 0.3) +
  labs(title = "Prior vs Posterior",
       x     = "Effect (log-odds scale)",
       y     = "Density") +
  xlim(c(-2, 6)) +
  theme_bw()

pp4 <- ggplot(draws_df) +
  geom_density(aes(`sd_ID__conditionlabelled:orientation0:aoi_typea:media_typea`), fill = "blue", alpha = 0.3) +
  geom_density(aes(`prior_sd_ID`), fill = "red", alpha = 0.3) +
  labs(title = "Prior vs Posterior",
       x     = "Effect (log-odds scale)",
       y     = "Density") +
  
  xlim(c(-2, 6)) +
  theme_bw()

pp5 <- ggplot(draws_df) +
  geom_density(aes(`sd_object_id__conditionlabelled:orientation0:aoi_typea:media_typea`), fill = "blue", alpha = 0.3) +
  geom_density(aes(`prior_sd_object_id`), fill = "red", alpha = 0.3) +
  labs(title = "Prior vs Posterior",
       x     = "Effect (log-odds scale)",
       y     = "Density") +
  xlim(c(-2, 6)) +
  theme_bw()

```


##visualise
```{r}
pp1

pp2

pp3

pp4

pp5

```





##  Hypothesis Testing


#################### Hypothesis 1a #############################################

```{r}
#H1a prop dwell time on art AOIs higher for 0 degree orientation on all objects compared to other orientations

hypothesis(m4_pos, "(conditionlabelled:orientation0:aoi_typea:media_typea + conditionunlabelled:orientation0:aoi_typea:media_typea +
    conditionlabelled:orientation0:aoi_typea:media_typet + conditionunlabelled:orientation0:aoi_typea:media_typet)/4 > (conditionlabelled:orientation90:aoi_typea:media_typea + conditionunlabelled:orientation90:aoi_typea:media_typea + conditionlabelled:orientation180:aoi_typea:media_typea + conditionunlabelled:orientation180:aoi_typea:media_typea + conditionlabelled:orientation270:aoi_typea:media_typea + conditionunlabelled:orientation270:aoi_typea:media_typea +
conditionlabelled:orientation90:aoi_typea:media_typet + conditionunlabelled:orientation90:aoi_typea:media_typet + conditionlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation180:aoi_typea:media_typet + conditionlabelled:orientation270:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/12")
```



################## Hypothesis 1b ##############################################


```{r}
 ##h1b
# tool will have higher trans_dwell in aoi_typet compared to aoi_typea 

#if it is a tool object, the dwell time will be higher in tool aois compared to art aois

hypothesis(m4_pos, "(conditionlabelled:orientation0:aoi_typet:media_typet +
  conditionlabelled:orientation90:aoi_typet:media_typet +
  conditionlabelled:orientation180:aoi_typet:media_typet +
  conditionlabelled:orientation270:aoi_typet:media_typet +
  conditionunlabelled:orientation0:aoi_typet:media_typet +
  conditionunlabelled:orientation90:aoi_typet:media_typet +
  conditionunlabelled:orientation180:aoi_typet:media_typet +
  conditionunlabelled:orientation270:aoi_typet:media_typet) / 8 >
  (conditionlabelled:orientation0:aoi_typea:media_typet +
  conditionlabelled:orientation90:aoi_typea:media_typet +
  conditionlabelled:orientation180:aoi_typea:media_typet +
  conditionlabelled:orientation270:aoi_typea:media_typet +
  conditionunlabelled:orientation0:aoi_typea:media_typet +
  conditionunlabelled:orientation90:aoi_typea:media_typet +
  conditionunlabelled:orientation180:aoi_typea:media_typet +
  conditionunlabelled:orientation270:aoi_typea:media_typet) / 8")
```

```{r}
### H1b --> 

# if the object is a tool, then the dwell time will be higher for tool aois compared to tool aois in art objects

hypothesis(m4_pos, "(conditionlabelled:orientation0:aoi_typet:media_typet +
  conditionlabelled:orientation90:aoi_typet:media_typet +
  conditionlabelled:orientation180:aoi_typet:media_typet +
  conditionlabelled:orientation270:aoi_typet:media_typet +
  conditionunlabelled:orientation0:aoi_typet:media_typet +
  conditionunlabelled:orientation90:aoi_typet:media_typet +
  conditionunlabelled:orientation180:aoi_typet:media_typet +
  conditionunlabelled:orientation270:aoi_typet:media_typet) / 8 >
  (conditionlabelled:orientation0:aoi_typet:media_typea +
  conditionlabelled:orientation90:aoi_typet:media_typea +
  conditionlabelled:orientation180:aoi_typet:media_typea +
  conditionlabelled:orientation270:aoi_typet:media_typea +
  conditionunlabelled:orientation0:aoi_typet:media_typea +
  conditionunlabelled:orientation90:aoi_typet:media_typea +
  conditionunlabelled:orientation180:aoi_typet:media_typea +
  conditionunlabelled:orientation270:aoi_typet:media_typea) / 8")


```

```{r}
###H1b 
#if the object is art then the dwell time will be higher in art aois than for tool aois


hypothesis(m4_pos, "(conditionlabelled:orientation0:aoi_typea:media_typea +
  conditionlabelled:orientation90:aoi_typea:media_typea +
  conditionlabelled:orientation180:aoi_typea:media_typea +
  conditionlabelled:orientation270:aoi_typea:media_typea +
  conditionunlabelled:orientation0:aoi_typea:media_typea +
  conditionunlabelled:orientation90:aoi_typea:media_typea +
  conditionunlabelled:orientation180:aoi_typea:media_typea +
  conditionunlabelled:orientation270:aoi_typea:media_typea) / 8 >
  (conditionlabelled:orientation0:aoi_typet:media_typea +
  conditionlabelled:orientation90:aoi_typet:media_typea +
  conditionlabelled:orientation180:aoi_typet:media_typea +
  conditionlabelled:orientation270:aoi_typet:media_typea +
  conditionunlabelled:orientation0:aoi_typet:media_typea +
  conditionunlabelled:orientation90:aoi_typet:media_typea +
  conditionunlabelled:orientation180:aoi_typet:media_typea +
  conditionunlabelled:orientation270:aoi_typet:media_typea) / 8")




```

```{r}
###H1b 

# If the object is art then the dwell time will be larger for art aois compared to art aois when the object is a tool

hypothesis(m4_pos, "(conditionlabelled:orientation0:aoi_typea:media_typea +
  conditionlabelled:orientation90:aoi_typea:media_typea +
  conditionlabelled:orientation180:aoi_typea:media_typea +
  conditionlabelled:orientation270:aoi_typea:media_typea +
  conditionunlabelled:orientation0:aoi_typea:media_typea +
  conditionunlabelled:orientation90:aoi_typea:media_typea +
  conditionunlabelled:orientation180:aoi_typea:media_typea +
  conditionunlabelled:orientation270:aoi_typea:media_typea) / 8 >
  (conditionlabelled:orientation0:aoi_typea:media_typet +
  conditionlabelled:orientation90:aoi_typea:media_typet +
  conditionlabelled:orientation180:aoi_typea:media_typet +
  conditionlabelled:orientation270:aoi_typea:media_typet +
  conditionunlabelled:orientation0:aoi_typea:media_typet +
  conditionunlabelled:orientation90:aoi_typea:media_typet +
  conditionunlabelled:orientation180:aoi_typea:media_typet +
  conditionunlabelled:orientation270:aoi_typea:media_typet) / 8")
```







######################## Hypothesis 2 ##########################################


```{r}
# H2
# tools will have higher relative trans_dwell for aoi_typet (compared to aoi_typea)
#when in labelled condition then in the unlabelled condition



hypothesis(m4_pos, "((conditionlabelled:orientation0:aoi_typet:media_typet - conditionlabelled:orientation0:aoi_typet:media_typea) +
  (conditionlabelled:orientation90:aoi_typet:media_typet - conditionlabelled:orientation90:aoi_typet:media_typea) +
  (conditionlabelled:orientation180:aoi_typet:media_typet - conditionlabelled:orientation180:aoi_typet:media_typea) +
  (conditionlabelled:orientation270:aoi_typet:media_typet - conditionlabelled:orientation270:aoi_typet:media_typea))/4 >
  ((conditionunlabelled:orientation0:aoi_typet:media_typet - conditionunlabelled:orientation0:aoi_typet:media_typea) +
  (conditionunlabelled:orientation90:aoi_typet:media_typet - conditionunlabelled:orientation90:aoi_typet:media_typea) +
  (conditionunlabelled:orientation180:aoi_typet:media_typet - conditionunlabelled:orientation180:aoi_typet:media_typea) +
  (conditionunlabelled:orientation270:aoi_typet:media_typet - conditionunlabelled:orientation270:aoi_typet:media_typea))/4")

```

```{r}
# H2 Do the same for art


# tool will have higher relative trans_dwell for aoi_typet (compared to aoi_typea)
#when in labelled condition then in the unlabelled condition


hypothesis(m4_pos, "((conditionlabelled:orientation0:aoi_typea:media_typea - conditionlabelled:orientation0:aoi_typea:media_typet) +
  (conditionlabelled:orientation90:aoi_typea:media_typea - conditionlabelled:orientation90:aoi_typea:media_typet) +
  (conditionlabelled:orientation180:aoi_typea:media_typea - conditionlabelled:orientation180:aoi_typea:media_typet) +
  (conditionlabelled:orientation270:aoi_typea:media_typea - conditionlabelled:orientation270:aoi_typea:media_typet))/4 >
  ((conditionunlabelled:orientation0:aoi_typea:media_typea - conditionunlabelled:orientation0:aoi_typea:media_typet) +
  (conditionunlabelled:orientation90:aoi_typea:media_typea - conditionunlabelled:orientation90:aoi_typea:media_typet) +
  (conditionunlabelled:orientation180:aoi_typea:media_typea - conditionunlabelled:orientation180:aoi_typea:media_typet) +
  (conditionunlabelled:orientation270:aoi_typea:media_typea - conditionunlabelled:orientation270:aoi_typea:media_typet))/4")



```



############################ Hypothesis 3 ######################################

```{r}
#H3 


hypothesis(m4_pos, "(conditionunlabelled:orientation0:aoi_typea:media_typea  + conditionunlabelled:orientation0:aoi_typea:media_typet)/2 - (conditionunlabelled:orientation90:aoi_typea:media_typea + conditionunlabelled:orientation180:aoi_typea:media_typea + conditionunlabelled:orientation270:aoi_typea:media_typea  + conditionunlabelled:orientation90:aoi_typea:media_typet + conditionunlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/6 >
  (conditionlabelled:orientation0:aoi_typea:media_typea  + conditionlabelled:orientation0:aoi_typea:media_typet)/2 - (conditionlabelled:orientation90:aoi_typea:media_typea + conditionlabelled:orientation180:aoi_typea:media_typea + conditionlabelled:orientation270:aoi_typea:media_typea  + conditionlabelled:orientation90:aoi_typea:media_typet + conditionlabelled:orientation180:aoi_typea:media_typet + conditionlabelled:orientation270:aoi_typea:media_typet)/6")
```









