---
title: "Model Comparison"
author: "Frederieke"
date: "2026-04-22"
output: html_document
---

--------------------------- Model Comparison -----------------------------------

This markdown provides an overview over the model comparison process for both models.

--------------------------------------------------------------------------------






```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




### Loading packages
```{r}
pacman::p_load(tidyverse, rlang, tidymodels, gridExtra, brms, cmdstanr, truncnorm, bayesplot, loo)
```



### Read in dataframe
```{r}
dt <- read_csv("C:/.../Analysis/all_data.csv")
```



################# Model 1 (H1a, H1b, H2, H3 ####################################



### Define models with varying complexity
```{r}

m1 <- bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type)

m2 <- bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type +
            (1|ID) +
            (1 | object_id))

m4 <- bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type + 
           (0 + condition:orientation:aoi_type:media_type | ID) + 
           (0 + condition:orientation:aoi_type:media_type | object_id))

```




######################### m1 ###################################################

```{r}
get_prior(
  bf(trans_dwell ~ 0 + condition:orientation:aoi_type:media_type,
     phi ~ 1),
  data   = dt,
  family = Beta()
)
```


```{r}
#mainly using weakly informative priors
pr1 <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor")
)
```



#### Run priors
```{r}
m1_prior <-
  brm(
    m1,
    data = dt,
    prior = pr1,
    family = Beta(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "m1_prior",
    warmup = 1000,
    iter = 2000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```



### Prior checks
```{r}
pp1 <- pp_check(m1_prior, ndraws = 100)
```

```{r}
pp1 + xlim(-2, 2)
```


### Fitting the posterior model
```{r}
m1_pos <- brm(
    m1,
    data = dt,
    family = Beta(link = "logit"),
    prior = pr1,
    sample_prior = T,
    iter = 2000,
    warmup = 1000,
    cores = 2,
    chains = 2,
    backend = "cmdstanr",
    threads = threading(4),
    file = "m1_pos",
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 15 ))


```


```{r}
summary(m1_pos)
```


## Check Posteriors
```{r}
pp1_pos <- pp_check(m1_pos, ndraws = 100) + xlim(-0.2, 1.2)
pp1_pos
```




###################### m2 ############################################


```{r}
#mainly using weakly informative priors
pr2 <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor")
)
```




```{r}
#### Run priors
m2_prior <-
  brm(
    m2,
    data = dt,
    prior = pr2,
    family = Beta(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "m2_prior",
    warmup = 1000,
    iter = 2000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```





### Prior checks
```{r}
pp2 <- pp_check(m2_prior, ndraws = 100)

pp2 + xlim(0,1.5)
```


### Fitting the posterior model
```{r}
m2_pos <- brm(
    m2,
    data = dt,
    family = Beta(link = "logit"),
    prior = pr2,
    sample_prior = T,
    iter = 2000,
    warmup = 1000,
    cores = 2,
    chains = 2,
    backend = "cmdstanr",
    threads = threading(4),
    file = "m2_pos",
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 15 ))


```



```{r}
summary(m2_pos)
```


```{r}
## Check Posteriors
pp2_pos <- pp_check(m2_pos, ndraws = 100) + xlim(-0.2, 1.2)
pp2_pos
```


######################## m4 ####################################################


```{r}
pr4 <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor")
)
```



#### Run priors
```{r}
m4_prior <-
  brm(
    m4,
    data = dt,
    prior = pr4,
    family = Beta(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "real_m4_prior",
    warmup = 500,
    iter = 1000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```



### Prior checks
```{r}
pp4 <- pp_check(m4_prior, ndraws = 100)
```

```{r}
pp4 + xlim(-1, 2)
```



### Fitting the posterior model
```{r}
m4_pos <- brm(
    m4,
    data = dt,
    family = Beta(link = "logit"),
    prior = pr4,
    sample_prior = T,
    iter = 2000,
    warmup = 2000,
    cores = 2,
    chains = 2,
    seed = 1012,
    backend = "cmdstanr",
    threads = threading(4),
    file = "real_m4_pos",
    control = list(
      adapt_delta = 0.95,
      max_treedepth = 15 ))


```


```{r}
summary(m4_pos)
```



#### Leave one out cross validation



```{r}
loo1 <- loo(m1_pos)

loo2 <- loo(m2_pos)

loo4 <- loo(m4_pos)
```


```{r}
loo_compare(loo1, loo2, loo4)

#4 performs best
```





####################### Model 2#####################################################


Model testing for hypothesis 4

#################################################################################



```{r}
dt <- dt %>% 
  rename(Classification = KeyPressEvent) #rename to classification
```


```{r}
# Recode key presses to meaningful labels and order them
dt <- dt %>%
  mutate(
    Classification = factor(
      case_when(
        Classification == "S" ~ "Tool",
        Classification == "C" ~ "Mostly Tool",
        Classification == "M" ~ "Mostly Art",
        Classification == "L" ~ "Art"
      ),
      levels  = c("Tool", "Mostly Tool", "Mostly Art", "Art"),
      ordered = TRUE
    )
  )

dt <- dt %>%
  filter(!is.na(Classification))
```





```{r}

dt_class <- dt %>%
  group_by(ID, MediaName, Classification,
           media_type, orientation, condition, object_id) %>%  # no aoi_type!
  summarise(
    prop_dwell_tool = sum(dwell_time_ms[aoi_type == "t"]) / 
                     sum(dwell_time_ms),
    prop_dwell_art  = sum(dwell_time_ms[aoi_type == "a"]) / 
                     sum(dwell_time_ms),
    .groups = "drop"
  )

```


#### Defining models
```{r}


m1_H4a <- bf(Classification ~ condition:media_type +
               condition:media_type:prop_dwell_tool)

m2_H4a <- bf(Classification ~ condition:media_type +
               condition:media_type:prop_dwell_art +
               (1|ID) + 
               (1|object_id))

m3_H4a <- bf(Classification ~ condition:media_type +
               condition:media_type:prop_dwell_tool +
    (condition:media_type + condition:media_type:prop_dwell_tool | ID) + 
      (condition + condition:prop_dwell_tool | object_id))

```







################# m1 ######################################################



#### set priors
```{r}
pr1_H4a <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor"),
  prior(normal(0, 3), class = "Intercept")

)
```



```{r}
#### Run priors
m1_prior_H4a <-
  brm(
    m1_H4a,
    data = dt_class,
    prior = pr1_H4a,
    family = cumulative(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "m1_prior_H4a",
    warmup = 1000,
    iter = 2000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```



### Prior checks
```{r}
pp1_H4a <- pp_check(m1_prior_H4a, ndraws = 100)
```

```{r}
pp1_H4a + xlim(0.5, 5)
```




### Fitting the posterior model
```{r}
m1_pos_H4a <- brm(
    m1_H4a,
    data = dt_class,
    family = cumulative(link = "logit"),
    prior = pr1_H4a,
    sample_prior = T,
    iter = 2000,
    warmup = 1000,
    cores = 2,
    chains = 2,
    seed = 1012,
    backend = "cmdstanr",
    threads = threading(4),
    file = "m1_pos_H4a",
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 15 ))


```

```{r}
summary(m1_pos_H4a)
```






################# m2 ######################################################



#### set priors
```{r}
pr2_H4a <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor"),
  prior(normal(0, 3), class = "Intercept")

)
```



```{r}
#### Run priors
m2_prior_H4a <-
  brm(
    m2_H4a,
    data = dt_class,
    prior = pr2_H4a,
    family = cumulative(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "m2_prior_H4a",
    warmup = 1000,
    iter = 2000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```



### Prior checks
```{r}
pp2_H4a <- pp_check(m2_prior_H4a, ndraws = 100)
```

```{r}
pp2_H4a + xlim(0.5, 5)
```




### Fitting the posterior model
```{r}
m2_pos_H4a <- brm(
    m2_H4a,
    data = dt_class,
    family = cumulative(link = "logit"),
    prior = pr2_H4a,
    sample_prior = T,
    iter = 2000,
    warmup = 1000,
    cores = 2,
    chains = 2,
    seed = 1012,
    backend = "cmdstanr",
    threads = threading(4),
    file = "m2_pos_H4a",
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 15 ))


```

```{r}
summary(m2_pos_H4a)
```





################# m3 ######################################################



#### set priors
```{r}
pr3_H4a <- c(
  prior(normal(0, 1.5), class = "b"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor"),
  prior(normal(0, 3), class = "Intercept")

)
```



```{r}
#### Run priors
m3_prior_H4a <-
  brm(
    m3_H4a,
    data = dt_class,
    prior = pr3_H4a,
    family = cumulative(link="logit"),
    sample_prior="only",
    backend = "cmdstanr",
    file = "m3_prior_H4a",
    warmup = 1000,
    iter = 2000,
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20
    )
  )
```



### Prior checks
```{r}
pp3_H4a <- pp_check(m3_prior_H4a, ndraws = 100)
```

```{r}
pp3_H4a + xlim(0.5, 5)
```




### Fitting the posterior model
```{r}
m3_pos_H4a <- brm(
    m3_H4a,
    data = dt_class,
    family = cumulative(link = "logit"),
    prior = pr3_H4a,
    sample_prior = T,
    iter = 2000,
    warmup = 1000,
    cores = 2,
    chains = 2,
    seed = 1012,
    backend = "cmdstanr",
    threads = threading(4),
    file = "m3_pos_H4a",
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 15 ))


```

```{r}
summary(m3_pos_H4a)
```




#### Leave one out cross validation



```{r}
H4_loo1 <- loo(m1_pos_H4a)

H4_loo2 <- loo(m2_pos_H4a)

H4_loo3 <- loo(m3_pos_H4a)
```


```{r}
loo_compare(H4_loo1, H4_loo2, H4_loo3)

#3 performs best
```








