---
title: "General Results Plots"
author: "Freddy"
date: "2026-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load packages
```{r}
pacman::p_load(tidyverse, gridEXTRA, ggtext, magick, png, extrafont, ggridges, brms, bayesplot)
```


```{r}
plot_df <- read_csv("C:/Users/frede/Documents/Speciale/Analysis/all_data.csv")
```


```{r}
#change key presses for more beautiful plots

plot_df <- plot_df %>%
  mutate(
    KeyPressEvent = factor(KeyPressEvent, levels = c("S","C","M","L")),
    media_type = factor(media_type, levels = c("t","a"),
                        labels = c("Tool","Art"))
  )

plot_df <- plot_df %>%
  mutate(
    response_type = recode(KeyPressEvent,
      "S" = "Tool",
      "C" = "Mostly Tool",
      "M" = "Mostly Art",
      "L" = "Art"
    )
  )

plot_df <- na.omit(plot_df)
```



```{r}
#order thos for plotting
plot_df$condition <- factor(
  plot_df$condition,
  levels = c("unlabelled", "labelled"),
  labels = c("Unlabelled", "Labelled")
)

plot_df$media_type <- factor(
  plot_df$media_type,
  levels = c("a", "t"),
  labels = c("Art Objects", "Tool Objects")
)


```










################## Classification Patterns - Plotting ##########################


```{r}
plot_df <- plot_df %>% 
  rename(Classification = KeyPressEvent)

# Recode key presses to meaningful labels and order them
plot_df <- plot_df %>%
  mutate(
    Classification = factor(
      case_when(
        Classification == "S" ~ "Tool",
        Classification == "C" ~ "Mostly Tool",
        Classification == "M" ~ "Mostly Art",
        Classification == "L" ~ "Art"
      ),
      levels  = c("Tool", "Mostly Tool", "Mostly Art", "Art"),
      ordered = TRUE
    )
  )

plot_df <- plot_df %>%
  filter(!is.na(Classification))
```




```{r}

class <- plot_df %>%
  count(condition, media_type, Classification) %>%
  group_by(condition, media_type) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = Classification, y = prop, fill = condition)) +
  geom_col(position = "dodge", alpha = 0.95) +
  scale_fill_manual(
    values = c("Unlabelled" = "#A1AF4F", "Labelled" = "#74346B"),
    name = "Condition"
  ) +
  facet_wrap(~media_type) +
  labs(
    x = "Classification Response",
    y = "Proportion of Response",
    title = "Classification by Object Type"
  ) +
  theme_bw() +
  theme(text = element_text(family = "Cambria", size = 22))

class

ggsave("responses.png", width = 14, height = 6)
```








########################## Boxplots of Dwell Time ##############################

```{r}
dwell_time_subset <- d_full %>%
  mutate(
    aoi_type_label   = ifelse(aoi_type == "t", "Tool AOI", "Art AOI"),
    media_type_label = ifelse(media_type == "t", "Tool Object", "Art Object"),
    rotation_label   = factor(paste0(media_rotation, "°"),
                              levels = c("0°", "90°", "180°", "270°"))
  )
    
dwell_time_subset <- na.omit(dwell_time_subset)
```




```{r}



# Filter to each condition
dwell_labelled <- dwell_time_subset %>%
  filter(condition == "labelled")

dwell_unlabelled <- dwell_time_subset %>%
  filter(condition == "unlabelled")

# Plot function to avoid repeating code
plot_dwell <- function(data, title) {
  ggplot(data, aes(x = aoi_type_label, y = prop_dwell, fill = aoi_type_label)) +
    geom_boxplot(outlier.shape = 23, alpha = 0.9) +
    facet_grid(media_type_label ~ rotation_label) +
    scale_fill_manual(values = c("Tool AOI" = "#AE4900", "Art AOI" = "#006B70")) +
    labs(
      title = title,
      x = "AOI",
      y = "Proportion of Dwell Time",
      fill  = "AOI Type"
    ) +
    theme_bw(base_size = 18) +
    theme(
    strip.text      = element_text(face = "bold", family = "Cambria", size = 14),
    axis.text       = element_text(family = "Cambria"),
    axis.title      = element_text(family = "Cambria"),
    legend.text     = element_text(family = "Cambria"),
    plot.title      = element_text(family = "Cambria"),
    legend.position = "bottom"
  )
}

# Generate plots
plot_labelled   <- plot_dwell(dwell_labelled,   "Labelled Condition")
plot_unlabelled <- plot_dwell(dwell_unlabelled, "Unlabelled Condition")


plot_labelled
plot_unlabelled

ggsave("boxplot_labelled.png",plot_labelled, width = 12, height = 8)
ggsave("boxplot_unlabelled.png", plot_unlabelled, width = 12, height = 8)
```






#################### Classification for results ################################

```{r}
# make new dataframe just in case
dt_new <- d_full

dt_new$Classification <- factor(
  dt_new$KeyPressEvent,
  levels = c("S", "C", "M", "L"),
  labels = c("Tool", "Mostly Tool", "Mostly Art", "Art")
)

dt_new$media_type <- factor(
  dt_new$media_type,
  levels = c("a", "t"),
  labels = c("Art Objects", "Tool Objects")
)
dt_new <- na.omit(dt_new)
```



```{r}
# Plot

dt_new %>%
  count(media_type, Classification) %>%
  group_by(media_type) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = Classification, y = prop, fill = Classification)) +
  geom_col(alpha = 0.9) +
  scale_fill_manual(values = c("#AE4900" , "#AE4900", "#006B70","#006B70")) +
  facet_wrap(~media_type) +
  labs(
    x = "Classification Response",
    y = "Proportion of Responses",
    title = "Classification Responses by Object Type"
  ) +
  theme_bw() +
  theme(text = element_text(family = "Cambria", size = 14))

ggsave("classification.png", width = 10, height = 6)
```



#### PLotting unlabelled and labelled individually
```{r}
df_un_class <- subset(dt_new, condition == "unlabelled")
df_la_class <- subset(dt_new, condition == "labelled")
```


```{r}
#unlabelled plot

un_class_plot <- df_un_class %>%
  count(media_type, Classification) %>%
  group_by(media_type) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = Classification, y = prop, fill = Classification)) +
  geom_col(alpha = 0.9) +
  scale_fill_manual(values = c("#AE4900" , "#AE4900", "#006B70","#006B70")) +
  facet_wrap(~media_type) +
  labs(
    x = "Classification Response",
    y = "Proportion of Responses",
    title = "Unlabelled - Classification Responses by Object Type"
  ) +
  theme_bw() +
  theme(text = element_text(family = "Cambria", size = 14))


#labelled plot
la_class_plot <- df_la_class %>%
  count(media_type, Classification) %>%
  group_by(media_type) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = Classification, y = prop, fill = Classification)) +
  geom_col(alpha = 0.9) +
  scale_fill_manual(values = c("#AE4900" , "#AE4900", "#006B70","#006B70")) +
  facet_wrap(~media_type) +
  labs(
    x = "Classification Response",
    y = "Proportion of Responses",
    title = "Labelled - Classification Responses by Object Type"
  ) +
  theme_bw() +
  theme(text = element_text(family = "Cambria", size = 14))


class_pl <- grid.arrange(un_class_plot, la_class_plot, nrow = 2)

class_pl

ggsave("classification.png", class_pl, width = 9, height = 10)
```










###################### Forest Plot #############################################

Making a bayesian forest plot, aka a posterior contrast plot, to highlight the 95-CI distributions of hypothesis results, coming from posterior draws.

################################################################################


#### 19:51
```{r}
eval_h_func <- function(h_string, draws_df) {
  
  # Splits on the inequality sign to get the difference
  if (grepl(">", h_string)) {
    parts <- strsplit(h_string, ">")[[1]]
  } else if (grepl("<", h_string)) {
    parts <- rev(strsplit(h_string, "<")[[1]])  # flip so it's still LHS - RHS
  } else {
    parts <- strsplit(h_string, "=")[[1]]
  }
  
  lhs <- trimws(parts[1])
  rhs <- trimws(parts[2])
  
  # Combine into a contrast: LHS - RHS
  expr_str <- paste0("(", lhs, ") - (", rhs, ")")
  
  col_names <- colnames(draws_df)
  col_names_sorted <- col_names[order(nchar(col_names), decreasing = TRUE)]
  
  for (col in col_names_sorted) {
    bare <- sub("^b_", "", col)
    pattern <- paste0("(?<!`)\\b", gsub("([.|()\\^{}+$*?])", "\\\\\\1", bare), "\\b")
    if (grepl(pattern, expr_str, perl = TRUE)) {
      expr_str <- gsub(pattern, paste0("`", col, "`"), expr_str, perl = TRUE)
    }
  }
  
  result <- tryCatch(
    with(draws_df, eval(parse(text = expr_str))),
    error = function(e) { message("Error in: ", h_string, "\n", e); return(NULL) }
  )
  return(result)
}
```


```{r}
eval_h_func <- function(h_string, draws_df) {
  
  if (grepl(">", h_string)) {
    parts <- strsplit(h_string, ">")[[1]]
  } else if (grepl("<", h_string)) {
    parts <- rev(strsplit(h_string, "<")[[1]])
  } else {
    parts <- strsplit(h_string, "=")[[1]]
  }
  
  lhs <- trimws(parts[1])
  rhs <- trimws(parts[2])
  expr_str <- paste0("(", lhs, ") - (", rhs, ")")
  
  col_names <- colnames(draws_df)
  col_names_sorted <- col_names[order(nchar(col_names), decreasing = TRUE)]
  
  for (col in col_names_sorted) {
    bare <- sub("^b_", "", col)
    pattern <- paste0("(?<!`)\\b", gsub("([.|()\\^{}+$*?])", "\\\\\\1", bare), "\\b")
    if (grepl(pattern, expr_str, perl = TRUE)) {
      expr_str <- gsub(pattern, paste0("`", col, "`"), expr_str, perl = TRUE)
    }
  }
  
  with(draws_df, eval(parse(text = expr_str)))
}
```


#### Define all hypotheses and labels 
```{r}
#hypotheses as list for input

hypotheses <- list(
  
  H1a = "(conditionlabelled:orientation0:aoi_typea:media_typea + conditionunlabelled:orientation0:aoi_typea:media_typea +
    conditionlabelled:orientation0:aoi_typea:media_typet + conditionunlabelled:orientation0:aoi_typea:media_typet)/4 >
    (conditionlabelled:orientation90:aoi_typea:media_typea + conditionunlabelled:orientation90:aoi_typea:media_typea +
    conditionlabelled:orientation180:aoi_typea:media_typea + conditionunlabelled:orientation180:aoi_typea:media_typea +
    conditionlabelled:orientation270:aoi_typea:media_typea + conditionunlabelled:orientation270:aoi_typea:media_typea +
    conditionlabelled:orientation90:aoi_typea:media_typet + conditionunlabelled:orientation90:aoi_typea:media_typet +
    conditionlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation180:aoi_typea:media_typet +
    conditionlabelled:orientation270:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/12",
  
    H1b_tool_art = "(conditionlabelled:orientation0:aoi_typet:media_typet + conditionlabelled:orientation90:aoi_typet:media_typet +
    conditionlabelled:orientation180:aoi_typet:media_typet + conditionlabelled:orientation270:aoi_typet:media_typet +
    conditionunlabelled:orientation0:aoi_typet:media_typet + conditionunlabelled:orientation90:aoi_typet:media_typet +
    conditionunlabelled:orientation180:aoi_typet:media_typet + conditionunlabelled:orientation270:aoi_typet:media_typet)/8 >
    (conditionlabelled:orientation0:aoi_typet:media_typea + conditionlabelled:orientation90:aoi_typet:media_typea +
    conditionlabelled:orientation180:aoi_typet:media_typea + conditionlabelled:orientation270:aoi_typet:media_typea +
    conditionunlabelled:orientation0:aoi_typet:media_typea + conditionunlabelled:orientation90:aoi_typet:media_typea +
    conditionunlabelled:orientation180:aoi_typet:media_typea + conditionunlabelled:orientation270:aoi_typet:media_typea)/8",
  
    H1b_art_tool = "(conditionlabelled:orientation0:aoi_typea:media_typea + conditionlabelled:orientation90:aoi_typea:media_typea +
    conditionlabelled:orientation180:aoi_typea:media_typea + conditionlabelled:orientation270:aoi_typea:media_typea +
    conditionunlabelled:orientation0:aoi_typea:media_typea + conditionunlabelled:orientation90:aoi_typea:media_typea +
    conditionunlabelled:orientation180:aoi_typea:media_typea + conditionunlabelled:orientation270:aoi_typea:media_typea)/8 >
    (conditionlabelled:orientation0:aoi_typea:media_typet + conditionlabelled:orientation90:aoi_typea:media_typet +
    conditionlabelled:orientation180:aoi_typea:media_typet + conditionlabelled:orientation270:aoi_typea:media_typet +
    conditionunlabelled:orientation0:aoi_typea:media_typet + conditionunlabelled:orientation90:aoi_typea:media_typet +
    conditionunlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/8",
  
  H1b_tool_aoi = "(conditionlabelled:orientation0:aoi_typet:media_typet + conditionlabelled:orientation90:aoi_typet:media_typet +
    conditionlabelled:orientation180:aoi_typet:media_typet + conditionlabelled:orientation270:aoi_typet:media_typet +
    conditionunlabelled:orientation0:aoi_typet:media_typet + conditionunlabelled:orientation90:aoi_typet:media_typet +
    conditionunlabelled:orientation180:aoi_typet:media_typet + conditionunlabelled:orientation270:aoi_typet:media_typet)/8 >
    (conditionlabelled:orientation0:aoi_typea:media_typet + conditionlabelled:orientation90:aoi_typea:media_typet +
    conditionlabelled:orientation180:aoi_typea:media_typet + conditionlabelled:orientation270:aoi_typea:media_typet +
    conditionunlabelled:orientation0:aoi_typea:media_typet + conditionunlabelled:orientation90:aoi_typea:media_typet +
    conditionunlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/8",
  
    H1b_art_aoi = "(conditionlabelled:orientation0:aoi_typea:media_typea + conditionlabelled:orientation90:aoi_typea:media_typea +
    conditionlabelled:orientation180:aoi_typea:media_typea + conditionlabelled:orientation270:aoi_typea:media_typea +
    conditionunlabelled:orientation0:aoi_typea:media_typea + conditionunlabelled:orientation90:aoi_typea:media_typea +
    conditionunlabelled:orientation180:aoi_typea:media_typea + conditionunlabelled:orientation270:aoi_typea:media_typea)/8 >
    (conditionlabelled:orientation0:aoi_typet:media_typea + conditionlabelled:orientation90:aoi_typet:media_typea +
    conditionlabelled:orientation180:aoi_typet:media_typea + conditionlabelled:orientation270:aoi_typet:media_typea +
    conditionunlabelled:orientation0:aoi_typet:media_typea + conditionunlabelled:orientation90:aoi_typet:media_typea +
    conditionunlabelled:orientation180:aoi_typet:media_typea + conditionunlabelled:orientation270:aoi_typet:media_typea)/8",
  
  
  H2_tool = "((conditionlabelled:orientation0:aoi_typet:media_typet - conditionlabelled:orientation0:aoi_typet:media_typea) +
    (conditionlabelled:orientation90:aoi_typet:media_typet - conditionlabelled:orientation90:aoi_typet:media_typea) +
    (conditionlabelled:orientation180:aoi_typet:media_typet - conditionlabelled:orientation180:aoi_typet:media_typea) +
    (conditionlabelled:orientation270:aoi_typet:media_typet - conditionlabelled:orientation270:aoi_typet:media_typea))/4 >
    ((conditionunlabelled:orientation0:aoi_typet:media_typet - conditionunlabelled:orientation0:aoi_typet:media_typea) +
    (conditionunlabelled:orientation90:aoi_typet:media_typet - conditionunlabelled:orientation90:aoi_typet:media_typea) +
    (conditionunlabelled:orientation180:aoi_typet:media_typet - conditionunlabelled:orientation180:aoi_typet:media_typea) +
    (conditionunlabelled:orientation270:aoi_typet:media_typet - conditionunlabelled:orientation270:aoi_typet:media_typea))/4",
  
  H3 = "(conditionunlabelled:orientation0:aoi_typea:media_typea + conditionunlabelled:orientation0:aoi_typea:media_typet)/2 -
    (conditionunlabelled:orientation90:aoi_typea:media_typea + conditionunlabelled:orientation180:aoi_typea:media_typea +
    conditionunlabelled:orientation270:aoi_typea:media_typea + conditionunlabelled:orientation90:aoi_typea:media_typet +
    conditionunlabelled:orientation180:aoi_typea:media_typet + conditionunlabelled:orientation270:aoi_typea:media_typet)/6 >
    (conditionlabelled:orientation0:aoi_typea:media_typea + conditionlabelled:orientation0:aoi_typea:media_typet)/2 -
    (conditionlabelled:orientation90:aoi_typea:media_typea + conditionlabelled:orientation180:aoi_typea:media_typea +
    conditionlabelled:orientation270:aoi_typea:media_typea + conditionlabelled:orientation90:aoi_typea:media_typet +
    conditionlabelled:orientation180:aoi_typea:media_typet + conditionlabelled:orientation270:aoi_typea:media_typet)/6",
  
  H4 = "(conditionlabelled:media_typea:prop_dwell_tool + conditionlabelled:media_typet:prop_dwell_tool +
    conditionunlabelled:media_typea:prop_dwell_tool + conditionunlabelled:media_typet:prop_dwell_tool)/4 > 0"
)

# Readable labels in the same order
hyp_labels <- c(
  "H1a: Canonical orientation on art AOI",
  "H1b.1: Tool AOI on tool vs. on art",
  "H1b.1: Art AOI on art vs. on tool",
  "H1b.2: Tool AOI focus on tools",
  "H1b.2: Art AOI focus on art",
  "H2: Labels bias visual attention",
  "H3: Labelling attenuates orientation",
  "H4: Dwell time predicts classification"
)
```


### Extracting draws from both models
```{r}

#merging draws from model m4_pos (H1a, H1b, H2, H3) and m3_pos_H4a (H4)

draws <- bind_cols(as_draws_df(m4_pos), as_draws_df(m3_pos_H4a))
# make sure column names dont clash
```


```{r}
# Hypotheses that belong to m4_pos
hypotheses_m4 <- hypotheses[c("H1a", "H1b_tool_art", 
                               "H1b_art_tool","H1b_tool_aoi", "H1b_art_aoi", "H2_tool", "H3")]

# Hypotheses that belong to m3_pos_H4a
hypotheses_m3 <- hypotheses["H4"]

# Run hypothesis on the correct model for each group
forest_data <- bind_rows(
  imap_dfr(hypotheses_m4, function(h, name) {
    res <- hypothesis(m4_pos, h)$hypothesis
    data.frame(
      label      = hyp_labels[which(names(hypotheses) == name)],
      estimate   = res$Estimate,
      ci_low     = res$CI.Lower,
      ci_high    = res$CI.Upper,
      evid_ratio = res$Evid.Ratio,
      supported  = res$Evid.Ratio >= 3
    )
  }),
  imap_dfr(hypotheses_m3, function(h, name) {
    res <- hypothesis(m3_pos_H4a, h)$hypothesis
    data.frame(
      label      = hyp_labels[which(names(hypotheses) == name)],
      estimate   = res$Estimate,
      ci_low     = res$CI.Lower,
      ci_high    = res$CI.Upper,
      evid_ratio = res$Evid.Ratio,
      supported  = res$Evid.Ratio >= 3
    )
  })
)
```


```{r}

#Extract posterior draws form all hypotheses

ridge_data <- imap_dfr(hypotheses, function(h, name) {
  vals <- eval_h_func(h, draws)
  if (is.null(vals)) return(NULL)
  data.frame(
    label = hyp_labels[which(names(hypotheses) == name)],
    value = vals
  )
})
```


```{r}
#fix order so H1a on top etc.

label_order <- rev(hyp_labels)   # rev() so H1a ends up on top
forest_data$label <- factor(forest_data$label, levels = label_order)
ridge_data$label  <- factor(ridge_data$label,  levels = label_order)
```

####Plotting a bayesian forest plot
```{r}
dwelling_in_a_forest <- ggplot() +
  geom_density_ridges(
    data  = ridge_data,
    aes(x = value, y = label),
    alpha = 0.8, linewidth = 0.5,
    color = "#454B21", fill = "#5D652D",
    trim  = FALSE, scale = 0.9   # scale < 1 keeps ridges from overlapping
  ) +
  geom_errorbar(
    data = forest_data,
    aes(y = label, xmin = ci_low, xmax = ci_high),
    width = 0, linewidth = 0.9, color = "#74346B"
  ) +
  geom_point(
    data  = forest_data,
    aes(y = label, x = estimate),
    fill  = "#74346B", color = "#74346B", size = 3
  ) +
  geom_vline(
    xintercept = 0, color = "#74346B",
    linetype = "dashed", linewidth = 0.8, alpha = 0.8 #line to emphasise 0 
  ) +
  labs(
    title = "Hypothesis Tests - Posterior Distributions",
    x     = "Contrast Estimate (95% CI)",
    y     = NULL
  ) +
  theme_bw() +
  scale_y_discrete(expand = expansion(add = c(0.3, 1))) +
  theme(
    text         = element_text(family = "Cambria", size = 16),
    legend.position = "bottom"
  )

dwelling_in_a_forest

ggsave("forestdwelll.png", dwelling_in_a_forest, width = 10, height = 6)
```










##################### Object Colllages #########################################

This section contains the code for building the collage figure, showcasing all 
40 objects used in the study.

################################################################################

```{r}

image_dir <- "C:/Users/frede/Documents/Speciale/Analysis/Objects/Art"


output_file <- "art_collage_border.png"

# Grid dimensions, 4 horizontal and 5 vertical
n_cols <- 4
n_rows <- 5

# Size in pixels!!!! to make sure they are identical
thumb_width  <- 200
thumb_height <- 120

# Spacing between images in pixels
gap <- 6

# Background colour blends in
bg_colour <- "black"



# load all images
files <- list.files(
  path       = image_dir,
  pattern    = "\\.png$",
  full.names = TRUE,
  ignore.case = TRUE
)


# Load and resize every image to a uniform thumbnail
thumbs <- lapply(files, function(f) {
  img <- image_read(f)
  image_resize(img, paste0(thumb_width, "x", thumb_height, "!"))
  # ! flag forces exact dimensions, looks weird otherwise

})


#Building collage

n_total <- n_cols * n_rows #should be 20


# Assemble rows and make collage
rows <- lapply(seq(1, n_total, by = n_cols), function(i) {
  row_imgs <- thumbs[i:(i + n_cols - 1)]
  image_append(image_join(row_imgs), stack = FALSE)
})

collage <- image_append(image_join(rows), stack = TRUE)

#save image
image_write(collage, path = output_file, format = "png")

```





```{r}
# Folder containing all tools
image_dir <- "C:/.../Analysis/Objects/Tool"

# Output file path and name
output_file <- "tool_collage_border.png"

# Grid dimensions, 4 horizontal and 5 vertical
n_cols <- 4
n_rows <- 5

# Size in pixels!!!! to make sure they are identical
thumb_width  <- 200
thumb_height <- 120

# Spacing between images in pixels
gap <- 6

# Background colour blends in
bg_colour <- "black"



# load all images
files <- list.files(
  path       = image_dir,
  pattern    = "\\.png$",
  full.names = TRUE,
  ignore.case = TRUE
)


# Load and resize every image to a uniform thumbnail
thumbs <- lapply(files, function(f) {
  img <- image_read(f)
  image_resize(img, paste0(thumb_width, "x", thumb_height, "!"))
  # ! flag forces exact dimensions, looks weird otherwise

})


#Building collage

n_total <- n_cols * n_rows #should be 20


# Assemble rows and make collage
rows <- lapply(seq(1, n_total, by = n_cols), function(i) {
  row_imgs <- thumbs[i:(i + n_cols - 1)]
  image_append(image_join(row_imgs), stack = FALSE)
})

collage <- image_append(image_join(rows), stack = TRUE)

#save image
image_write(collage, path = output_file, format = "png")

```












########################## Heatmaps ############################################

Making comparison heatmaps for both the unlabelled and the labelled group

################################################################################

```{r}
#separate the conditions
dun <- d_full %>% 
  filter(condition == "unlabelled")

dla <- d_full %>% 
  filter(condition == "labelled")

```





######################### Unlabelled Heatmaps ##################################

#0
```{r}
img <- readPNG("C:/Users/frede/Documents/Speciale/Analysis/Objects/Tool/27_t_0.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
          
# flip Y in the data itself instead of the axis
fixations_plot <- dun %>%
  filter(MediaName == "27_t_0.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_un_27_t_0.png", width = 6.5, height = 4)

```

#90

```{r}
img <- readPNG("C:/Users/frede/Documents/Speciale/Analysis/Objects/Art/30_a_90.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
# flip Y in the data itself instead of the axis
fixations_plot <- dun %>%
  filter(MediaName == "30_a_90.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_un_30_a_90.png", width = 6.5, height = 4)

```




### 03_t_180
```{r}
img <- readPNG("C:/Users/frede/Documents/Speciale/Analysis/Objects/Tool/03_t_180.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
# flip Y in the data itself instead of the axis
fixations_plot <- dun %>%
  filter(MediaName == "03_t_180.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_un_03_t_180.png", width = 6.5, height = 4)

```





#270
```{r}
img <- readPNG("C:/Users/frede/Documents/Speciale/Analysis/Objects/Art/06_a_270.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
          
# flip Y in the data itself instead of the axis
fixations_plot <- dun %>%
  filter(MediaName == "06_a_270.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_un_06_a_270.png", width = 6.5, height = 4)

```






########################## Labelled heatmaps ##################################

Example heatmaps of four different objects, two tools and two art objects, in four different orientations. Repeat entire process for labelled and unlabelled group, for side-by-side comparison.

################################################################################

#0
```{r}
img <- readPNG("C:/...e/Analysis/Objects/Tool/27_t_0_l.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
          
# flip Y in the data itself instead of the axis
fixations_plot <- dla %>%
  filter(MediaName == "27_t_0.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_la_27_t_0.png", width = 6.5, height = 4)

```





#90

```{r}
img <- readPNG("C:/.../Analysis/Objects/Art/30_a_90_l.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
# flip Y in the data itself instead of the axis
fixations_plot <- dla %>%
  filter(MediaName == "30_a_90.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_la_30_a_90.png", width = 6.5, height = 4)

```






### 03_t_180
```{r}
img <- readPNG("C:/.../Analysis/Objects/Tool/03_t_180_l.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
# flip Y in the data itself instead of the axis
fixations_plot <- dla %>%
  filter(MediaName == "03_t_180.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density") +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "right",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria"),
                              legend.text = element_text(family = "Cambria")
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_la_03_t_180.png", width = 6.5, height = 4)

```







#270
```{r}
img <- readPNG("C:/.../Analysis/Objects/Art/06_a_270_l.png")
height <- dim(img)[1]  # rows = height
width  <- dim(img)[2]  # cols = width
img_raster <- as.raster(img)
```


```{r}
          
# flip Y in the data itself instead of the axis
fixations_plot <- dla %>%
  filter(MediaName == "06_a_270.png") %>%
  mutate(y_flipped = height - `GazePointY (MCSpx)`)

ggplot(fixations_plot, aes(x = `GazePointX (MCSpx)`, y = y_flipped)) +
  annotation_raster(img_raster, xmin=0, xmax=width, ymin=0, ymax=height) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon",
                  contour = TRUE, alpha = 0.5,
                  h = c(50, 50)) +
  scale_fill_gradient(low = "#006B70", high = "#923F00", name = "Fixation Density",limits = c(0, 1),
  breaks = c(0.05, 0.93), labels = c("Low", "High")) +
  scale_x_continuous(limits = c(0, width), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, height), expand = c(0,0)) +
  coord_fixed() +
  theme_bw() +
  theme(
  axis.text = element_blank(),
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  legend.position = "bottom",
  plot.background = element_rect(fill = 'black', colour = 'black'),
                              legend.title = element_text(family = "Cambria", size = 12),
                              legend.text = element_text(family = "Cambria", size = 10)
)
  #theme(legend.position = "right") +
  #theme(plot.background = element_rect(fill = 'black', colour = 'black'))

ggsave("hm_la_06_a_270.png", width = 14, height = 8)

```













###################### General Results Overview ################################

This section contains some general calculations covering the results, such as demographics within the participant population, recording duration, as well as overall classification results.

################################################################################



########################## Average Gaze Duration ##################

```{r}
## keep only one value for each participant to calculate the mean time, value repeats because multiple rows
df_duration <- d_full %>% 
  group_by(ParticipantName, RecordingDuration) %>% 
  slice_head() 

mean(df_duration$RecordingDuration)
sd(df_duration$RecordingDuration)


df_duration_u <- df_duration %>% 
  filter(condition == "unlabelled")
mean(df_duration_u$RecordingDuration)
sd(df_duration_u$RecordingDuration)


df_duration_l <- df_duration %>% 
  filter(condition == "labelled")
mean(df_duration_l$RecordingDuration)
sd(df_duration_l$RecordingDuration)
```







#### Observations about dwell time
```{r}
dwell_time_subset %>%
  group_by(media_type_label, aoi_type_label) %>%
  summarise(
    min    = min(prop_dwell),
    max    = max(prop_dwell),
    q25    = quantile(prop_dwell, 0.25),
    q75    = quantile(prop_dwell, 0.75),
    IQR    = IQR(prop_dwell),
    .groups = "drop"
  )
```



### All classifications
```{r}
table(unlist(dt$KeyPressEvent))


df_un <- subset(dt, condition == "unlabelled")
df_l <- subset(dt, condition == "labelled")

```


### Unlabelled Classifications
```{r}
#14.003 --> All keypresses

table(unlist(df_un$KeyPressEvent))

#7069
```


### Labelled Classifications
```{r}
table(unlist(df_l$KeyPressEvent))
#6934
```






################## Demographic analysis #######################################


```{r}

d_dem <- d_full %>%
  distinct(ParticipantName, `[Gender]Value`, `[Age]Value`)

table(participants$`[Age]Value`)

# weirdly gone from the data, so added manually
participants <- d_dem %>%
  add_row(
    ParticipantName = "P45U1",
    '[Gender]Value' = "Female",
    '[Age]Value' = "25-34"
  )

```



```{r}
participants <- rbind(participants, data.frame(
  ParticipantName = "P45U1",
  `[Gender]Value` = "Female",
  `[Age]Value` = "25-34"
))

participants$age_midpoint <- dplyr::case_when(
  participants$`[Age]Value`== "18-24" ~ 21,
  participants$`[Age]Value` == "25-34" ~ 29.5,
  participants$`[Age]Value` == "35-44" ~ 39.5,
  participants$`[Age]Value` == "45-54" ~ 49.5,
  participants$`[Age]Value` == "65+" ~ 70
)

mean(participants$age_midpoint, na.rm = TRUE)
sd(participants$age_midpoint, na.rm = TRUE)
```



