---
title: "RETP Pre-processing"
author: "Frederieke"
date: "2026-03-18"
---


---------------------- Pre-processing pipeline ---------------------------------

This document includes the entire data pre-processing pipeline, preparing the raw eye-tracking data for the statistical analysis.

--------------------------------------------------------------------------------



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load packages
```{r}
pacman::p_load(readr, tidyverse, eyetrackingR, dplyr, sr, stringr, jpeg, grid, vroom, roperators)
```


### Read in individual data files, as the amount of data is too large for a single file

```{r}
df <- read_tsv("C:.../Analysis/retp4.tsv")
```




### See, which participants are part of the specific data file
```{r}
unique(df$ParticipantName)

unique(df$KeyPressEvent) #see if any false key presses
```



#### Remove last participant as double in dataframes
```{r}
#remove the last participant, because some show up both at the end and at the beginning of the next data file
df <- df[df$ParticipantName != "Test", ]

```


### Extract classification
```{r}
#remove key presses, which are errors
#rename key presses, which next to original ones when clear which key, to correct errors

classifications <- df %>%
  filter(!is.na(KeyPressEvent) & KeyPressEvent != "") %>%
  distinct(ParticipantName, MediaName, KeyPressEvent) %>%
  filter(KeyPressEvent != "Space") %>%
  filter(KeyPressEvent != "Return") %>%
  filter(KeyPressEvent != "Right") %>% 
  filter(KeyPressEvent != "NA") %>% 
  mutate(KeyPressEvent = case_when(
    KeyPressEvent %in% c("W", "D", "A") ~ "S",
    TRUE ~ KeyPressEvent)) %>% 
  mutate(KeyPressEvent = case_when(
    KeyPressEvent %in% c("X", "V") ~ "C",
    TRUE ~ KeyPressEvent)) %>% 
  mutate(KeyPressEvent = case_when(
    KeyPressEvent %in% c("Oemtilde", "K") ~ "L",
    TRUE ~ KeyPressEvent))
```



#### Keep only fixations
```{r}
fixations <- df %>%
  filter(GazeEventType == "Fixation")
```



####  Remove invalid gaze
```{r}
# Tobii validity: 0 = perfect, 1 = good, 2-4 = unreliable
fixations <- fixations %>%
  filter(ValidityLeft <= 1 | ValidityRight <= 1)
```


####  Keep one unique row per fixation
```{r}
#condense data frame
fixations <- fixations %>%
  distinct(ParticipantName, MediaName, FixationIndex, .keep_all = TRUE)
```



```{r}
# see if AOI names have dots instead of brackets? Then change next chunk if necessary, usually brackets
colnames(fixations) %>% str_subset("AOI")
```


### 4. Pivot AOI columns to long format
```{r}
#find proper AOI names to keep
# excludes secondary AOIs ending in _s before the closing bracket
aoi_cols <- colnames(fixations) %>%
  str_subset("^AOI\\[\\d+_[ta]_") %>%
  str_subset("_s", negate = TRUE) 

fixations_long <- fixations %>%
  select(ParticipantName, MediaName, FixationIndex, GazeEventDuration, `GazePointX (MCSpx)`, `GazePointY (MCSpx)`, KeyPressEvent, `[Gender]Value`, `[Age]Value`,RecordingDuration, all_of(aoi_cols)) %>%
  pivot_longer(
    cols = all_of(aoi_cols),
    names_to = "AOI_raw",
    values_to = "Hit"
  ) %>%
  filter(Hit == 1)  #fixations with no aoi hits excluded
```


#### Parse AOI names
```{r}

fixations_long <- fixations_long %>%
  mutate(
    AOI_clean   = str_extract(AOI_raw, "(?<=\\[)[^\\]]+"),  # get content inside, change to "(?<=AOI\\.).*(?=\\.Hit)" when dots instead of brackets
    object_id   = str_extract(MediaName, "^\\d+"), #first digits from media name
    aoi_type    = str_extract(AOI_clean, "(?<=_)[ta](?=_)"), #either t or a
    orientation    = str_extract(MediaName, "\\d+(?=\\.(png|png$)"), #
    aoi_index   = str_extract(AOI_clean, "\\d+$")
  )



```


### Parse media-name for cross-checking
```{r}
#verify that abiove works
fixations_long <- fixations_long %>%
  mutate(
    media_object   = str_extract(MediaName, "^\\d+"),
    media_type     = str_extract(MediaName, "(?<=_)[ta](?=_)"),
    orientation = str_extract(MediaName, "\\d+(?=\\.png$)")
  )
```


#### Sanity checks an overview
```{r}
# How many fixations per participant?
fixations_long %>%
  count(ParticipantName)

# How many fixations hit tool vs art AOI?
fixations_long %>%
  count(aoi_type)

#AOI parsing looks right?
fixations_long %>%
  select(AOI_raw, AOI_clean, object_id, aoi_type, orientation, aoi_index) %>%
  distinct() %>%
  head(20)
```



```{r}
fixations_long <- rename(fixations_long, ID = ParticipantName) #forgot to rename ID for simplicity
fixations_long <- fixations_long %>% 
  rename(Classification = KeyPressEvent) #and changed my mind about renaming this one
```


#### Summing gaze event duration per participant, stimulus, AOI
```{r}
 
dwell_time <- fixations_long %>%
  group_by(ID, MediaName, object_id, media_type, orientation, aoi_type, `[Gender]Value`, `[Age]Value`, RecordingDuration, `GazePointX (MCSpx)`, `GazePointY (MCSpx)` ) %>%
  summarise(
    dwell_time_ms    = sum(GazeEventDuration, na.rm = TRUE),
    n_fixations      = n(),
    .groups = "drop"
  )

#GazePoints on media for heatmaps
```


### Transform to proportion of dwell time
```{r}
 
dwell_time <- dwell_time %>%
  group_by(ID, MediaName) %>%
  mutate(
    total_dwell_ms   = sum(dwell_time_ms),
    prop_dwell       = dwell_time_ms / total_dwell_ms
  ) %>%
  ungroup()
```



#### Adding participant groups as column
```{r}

dwell_time <- dwell_time %>%
  mutate(
    condition = case_when(
      str_detect(ID, "L") ~ "labelled",
      str_detect(ID, "U") ~ "unlabelled",
      TRUE ~ NA_character_
    )
  )
```

####merge classifications
```{r}
dwell_time <- dwell_time %>%
  left_join(classifications, by = c("ID", "MediaName"))
```




### save as csv
```{r}
write_csv(dwell_time, "C:.../Analysis/P04L2P60L2.csv")
```



Repeat for all five data files


-------------------------- Combine all datafiles -------------------------------



#### Read in data
```{r}
d1 <- read_csv("C:/.../Analysis/P05U1P41U1.csv") 

d2 <- read_csv("C:/.../Analysis/P45U1P22L1.csv")

d3 <- read_csv("C:/.../Analysis/P26L1P15U2.csv") 

d4 <- read_csv("C:/.../Analysis/P19U2P61U2.csv") 

d5 <- read_csv("C:/.../Analysis/P04L2P60L2.csv")
```





### Merge dataframe

```{r}
d_full <- rbind(d1, d2, d3, d4, d5)
```

```{r}
write_csv(d_full, "C:/Users/frede/Documents/Speciale/Analysis/all_data.csv")
```

```{r}
unique(d_full$ParticipantName) #are all participants there?
```



### Data transformation

```{r}
#transforming the data to exclude 0's & 1's, avoiding singularity
d_full$trans_dwell <- ((d_full$prop_dwell * (d_full$total_dwell_ms - 1)) + 0.5) / d_full$total_dwell_ms

```


```{r}
# Set factor levels with 0 as baseline
d_full$orientation <- factor(d_full$orientation, 
                              levels = c("0", "90", "180", "270"))

# Double check
levels(d_full$orientation)

```








